Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-12-19 manuscript.tex
second revisions

In addressing all of these questions, it has been \change{important} to obtain
estimates of \emph{implicature rates}. For 1., implicature rates from
experimental tasks can be taken to inform whether scalar implicatures
should be considered default inferences. For 2., processing measures on
responses that indicate implicatures can be compared to processing
measures on responses that indicate literal interpretations. For 3.,
contextual effects can be examined by comparing implicature rates across
contexts. For 4., implicature rates can be compared across scales (or
across implicature types). For 5., implicature rates can be compared
across age groups.

@@

\change{Using the proportion of \enquote{FALSE} responses on true but underinformative trials as a proxy for implicature rate is  common in experimental pragmatics. For example, in one of the first studies to investigate scalar implicatures experimentally, Noveck (2001) tested adults' and children's interpretations of the scalar items \emph{might} and \emph{some}. The dependent measure in Noveck (2001) was the rate of ``logically correct responses'', i.e. responding ``yes'' to statements such as \emph{Some giraffes have long necks} or \emph{There might be a parrot [in the box]} when there had to be a parrot in the box. He found that children responded ``yes'' more frequently than adults, and concluded that children interpret scalar items \emph{some} and \emph{might} more logically (i.e. literally). Similarly in another landmark study, Papafragou and Musolino (2003) tested children and adults interpretation of the following set of scalar items: <\emph{two}, \emph{three}>, <\emph{some}, \emph{all}>, and <\emph{finish}, \emph{start}>. The dependent measure in this study was the proportion of ``No'' responses to a puppet's underinformative] statement. The study concluded that ``while adults overwhelmingly rejected infelicitous descriptions, children almost never did so''. Furthermore, the study compared  implicature rates across scales and concluded that ``children also differed from from adults in that their rejection rate on the numerical scale was reliably higher than on the two other scales.'' In their final experiment, Papafragou and Musolino (2003) modified their task to invite scalar inferences more easily. They reported that this manipulation resulted in a significantly higher rejection rates. Based on these results, they concluded that children's ability to compute implicatures is affected by the type of scalar item as well as children's awareness of the task's goals. Since these early pioneering studies, the rate of ``FALSE'' (or ``No'', ``Wrong'', ``Disagree'') responses on underinformative trials in truth-value judgment tasks has become a commonplace dependent measure (Doran et al (2012),  Geurts and Pouscoulous (2009), Potts, Lassiter, Levy, and Frank (2016), inter alia.)}

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of ``FALSE' responses in truth value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011); \change{Sikos et al (this issue)}). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute \change{a small, a big, or a huge strawberry } to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average \change{smaller} strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

@@

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
\change{assumes} two things: first, that whatever participants are doing in a
truth value judgment task, the behavioral measure can be interpreted as
providing a measure of interpretation; and second, that listeners either
do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues.
Following Degen and Goodman (2014), we will offer some remarks on why
truth value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking hypotheses.
We then discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking hypotheses in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a
depiction of either one or two animals (see Figure
\ref{fig:linkvisualization} for an example). As in a standard truth
value judgment task, the listener then observes an underinformative
utterance about this card (e.g., \enquote{There is a cat or a dog on the
card}) and is asked to provide a judgment on a scale with 2, 3, 4, or 5
response options, with endpoints \enquote{wrong} and
\enquote{right.}\footnote{An open question concerns the extent to which
  the labeling of points on the scale affects judgments (e.g.,
  \enquote{wrong}--\enquote{right} vs. \enquote{false}--\enquote{true}
  vs. \enquote{disagree}--\enquote{agree}). \change{Studies vary in the labeling of scale points.}} In the binary case, this reproduces the standard
truth value judgment task. Figure \ref{fig:linkvisualization}
exemplifies (some of) the researcher's options for grouping responses.
Under what we will call the \enquote{Strong link} assumption, only the
negative endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Under the \enquote{Weak link}
assumption, in contrast, any response that does not correspond to the
positive endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Intermediate grouping schemes are also
possible, but these are the ones we will consider here. Note that for
the binary case, the Weak and Strong link return the same categorization
scheme, but for any number of response options greater than 2, the Weak
and Strong link can in principle lead to differences in inferences about
implicature rate.

@@

Our primary goal in this study was to test whether the estimated
implicature rate in the experimental task is affected by the linking
hypothesis and the number of response options available to participants.
To this end, we only analyzed the critical trials (exhaustive and
scalar). In particular, we classified each data point from critical
trials as constituting an implicature (1) or not (0) under the Strong
and Weak link. Figure \ref{fig:implicatureRatePlot} shows the resulting
implicature rates by condition and link. \change{It is immediately apparent that there is variability in inferred implicature rate. In particular, the Weak link appears to result in greater estimates of
implicature rates in tasks with four or five response options, compared to the Strong link. For the binary and ternary task, the assumed link appears to play a much smaller role.}

@@

The model provided evidence for the following effects: First, there was
a main effect of trial type such that scalar trials resulted in greater
implicature rates than exhaustive trials (Mean Estimate = 6.09, 95\%
Credible Interval={[}1, 12.29{]}). Second, there was an interaction
between link and number of response options such that the quaternary
task (Mean Estimate = 14.03, 95\% Credible Interval={[}7.24, 21.88{]})
and the quinary task (Mean Estimate = 17.28, 95\% Credible
Interval={[}10.64, 25.80{]}) resulted in greater
implicature rates \change{with a weak link than with a strong link, but there was no evidence of a link-dependent difference in inferred implicature rate for the binary and ternary task}. Finally, there was a three-way interaction between
link, trial type, and number of response options, \change{driven by the binary/quinary contrast} (Mean Estimate = -7.74,
95\% Credible Interval={[}-16.59, -0.16{]}). \change{Simple effects analysis on only the binary and and quinary trials, separately for the exhaustive and scalar subset of the data, revealed that the three-way interaction is driven by a different effect of number of response options under the Weak vs Strong link for the two inference types. Specifically, on exhaustive trials, number of response options (2 vs.~5) only resulted in greater implicature rates under the Weak ($\beta$ = .2, $p <$ .0001), but not the Strong link ($\beta$ = -.8, $p <$ .82). In contrast, on scalar trials, number of response options (2 vs.~5) resulted in greater implicature rates under the Weak ($\beta$ = 3.6, $p <$ .005) link, and in lower implicature rates under the Strong link ($\beta$ = -4.0, $p <$ .0007).}

\change{In sum,} both number of response options and link
affected the inferred implicature rate\change{, as did the type of inference (exhaustive vs.~scalar)}.

@@

We take inspiration from the RSA approach and treat participants'
behavior in our experimental tasks as the result of a soft-optimal
pragmatic speaker in the RSA framework. That is, following Degen and
Goodman (2014), we proceed on the assumption that behavior on sentence
verification tasks such as truth value judgment tasks, is best modeled
as a function of an individual's mental representation of a cooperative
\change{speaker} (\(S_1\) in the language of RSA) rather than of a pragmatic
listener who interprets utterances (\(P_{L_1}\)).\footnote{\change{Degen and Goodman (2014) argue that sentence verification is more plausibly construed as a production task rather than as an interpretation task because participants, unlike in natural language comprehension, are provided with the ground truth about the state of the world that a speaker is describing. Thus, participants are in essence being asked to assess the quality of a speaker's utterance. In contrast, Degen and Goodman argue, true interpretation tasks are characterized by the listener inferring what the state of the world is that the speaker is describing, for instance by selecting from one of multiple interpretation options.}} In their paper, Degen
\& Goodman \change{show} that sentence verification tasks are relatively more
sensitive to contextual \change{features like} the
Question Under Discussion than are sentence interpretation tasks, and
that this follows if sentence interpretation tasks -- but not sentence
verification tasks -- require an additional layer of counterfactual
reasoning about the intentions of a cooperative speaker.

@@

The linking hypothesis under discussion assumes that speaker
probabilities of utterance given meaning are invariant across a) our
four different experimental conditions, b) across participants, and c)
within participants (that is, participants do not update
their \(S_1\) distribution in a local discourse context). We note that
the assumption (b) may conceivably be relaxed by allowing one or more of
the parameters in the model -- including the prior probability over
world states \(P_w\), the cost function on utterances \(C\), or the
rationality parameter \(\alpha\) -- to vary across participants. We also
note that assumption (c) in particular is in tension with a growing body
of empirical evidence that semantic and pragmatic interpretation is
modulated by rapid adaptation to the linguistic and social features of
one's interlocutors (Fine, Jaeger, Farmer, \& Qian, 2013; Kleinschmidt
\& Jaeger, 2015; \change{Yildirim, Degen, Tanenhaus, \& Jaeger, 2016)}.

However, if we should like to keep the above \change{simplifying} assumptions in place, then
this linking hypothesis commits us to explaining variation in the data in terms
of the threshold parameters of our responder model \(R\). Consider first
the variation in response across different experimental conditions on a
given trial, e.g.~evaluation of a guess of \enquote{cat and dog} when
the card contains both a cat and a dog. The variation in the proportion
of responses of \enquote{right} on this trial between the binary,
ternary, quaternary, and quinary conditions indicates that the threshold
value for \enquote{right} responses must vary across conditions; that
is, we predict that the \(\theta\) of the binary condition will differ
from, e.g., the \(\theta_1\) of the ternary condition as well as the
\(\theta_1\) of the quaternary condition. We also observed variation in
response on this trial within a single condition (for example, a
sizeable minority of participants responded \enquote{wrong} to this
trial in the binary condition). Thus, this linking hypothesis is
committed to the notion that threshold values may vary across
participants, such that a speaker probability of utterance \(S_1\)(u
\textbar{} s) can fall below \(\theta\) for some subset of participants
while \(S_1\)(u \textbar{} s) itself remains constant across
participants.

Lastly, for two utterances of the same
conditional probability and in the same experimental condition, 
participants in our experiment sometimes provided a judgment of
\enquote{right} to one utterance but \enquote{wrong} to the other. That
is, there was within-subject variation in this
experiment. One way to represent such variation would be to posit that
the parameterization of threshold values proceeds stochastically and
that threshold values are recalibrated for every individual sentence
verification task. Rather than representing a threshold as a discrete
value N between 0 and 1, we can represent that threshold as a
distribution over possible threshold values -- with mass centered around
N. Whenever an individual encounters  a single trial of our truth value judgment task experiment, a
threshold value is sampled from this distribution. \change{By  allowing values of \(\theta\) to vary stochastically in this way, we can capture that \(S_1\)(u \textbar{} s) can fall both above and below \(\theta\) for a given
participant.}

\change{The model in its present form already captures an interesting asymmetry in inferred implicature rates between exhaustive and scalar trials of the experiment: note specifically (c.f. Figure 8) that inferred implicature rates are greater in the binary and ternary conditions for scalar trials over exhaustive trials. This is expected given the model's inferred speaker probabilities: the speaker probability of producing ``There is a cat on the card'' in the context of there being a cat and dog on the card (an exhaustive implicature-inducing trial) is greater than the speaker probability of producing ``There is a cat or a dog on the card'' in that same context (a scalar implicature-inducing trial). Assuming noisy  $\theta$ values centered around N, participants are expected to respond `Right' more frequently on exhaustive than on scalar trials, which is precisely what is observed. Recall that these probabilities were derived via the simplifying assumption of uniform cost on utterances; in fact, adding cost to relatively complex disjunctive sentences over simple declarative sentences only predicts a more pronounced asymmetry in the experimentally-observed direction.}

@@

For the time being, however, we present the above analysis as a proof of
concept for the following idea: by relaxing the assumptions of the
traditional view of scalar implicature---namely, that scalar implicatures
either are or are not calculated, and that behavior on sentence
verification tasks directly reflects this binary interpretation
process---we can propose quantitative models of the variation in
behavior that is observed in experimental settings. We note that the linking
hypothesis proposed here is just one in the space of possible hypotheses. For
example, one might reject this threshold-based analysis in favor of one
whereby responses are the outcomes of sampling on the (pragmatic speaker
or pragmatic listener) probability distributions provided by an RSA
model. We \change{leave this systematic, quantitative} investigation to future work. For now we
emphasize that \change{explicit computational modeling of behavioral responses is a tool that is  available to researchers in experimental
pragmatics. While using the RSA framework as the modeling tool requires revising traditional assumptions about the nature of scalar
implicature by relaxing the crisp notion of scalar
implicature as something that is or is not \enquote{calculated} in
interpretation, it provides new flexibility to explicitly discuss
 behavior in experimental settings. One need not adopt the RSA framework as the tool for hypothesizing and testing the link between theoretical constructs and behavior in pragmatic experiments. However, the empirical findings we have reported here---that the inferences researchers draw about ``implicature rate'' are volatile and depend on various features of the paradigm and the linking hypothesis employed--- strongly suggest that experimental pragmatics  as a field must engage more seriously with the foundational questions of what we are measuring in the experiments we run.}

@@

\change{\hypertarget{ref-Sikos2019}{}
Sikos, L., Kim, M., \& Grodner, D.J. (in press). Social context modulates tolerance for pragmatic violations in binary but not graded judgments. \emph{Frontiers in Psychology}.}

@@

\change{\hypertarget{ref-Yildirim2016}{}
Yildirim, I., Degen, J., Tanenhaus, M.K., \& Jaeger, T.F. (2016). Talker-specificity and adaptation in quantifier interpretation. \emph{Journal of Memory and Language}, \emph{87}, pp. 128-143.}

>>>>

In addressing all of these questions, it has been important to obtain
estimates of \emph{implicature rates}. For 1., implicature rates from
experimental tasks can be taken to inform whether scalar implicatures
should be considered default inferences. For 2., processing measures on
responses that indicate implicatures can be compared to processing
measures on responses that indicate literal interpretations. For 3.,
contextual effects can be examined by comparing implicature rates across
contexts. For 4., implicature rates can be compared across scales (or
across implicature types). For 5., implicature rates can be compared
across age groups.

@@

Using the proportion of \enquote{FALSE} responses on true but underinformative trials as a proxy for implicature rate is  common in experimental pragmatics. For example, in one of the first studies to investigate scalar implicatures experimentally, Noveck (2001) tested adults' and children's interpretations of the scalar items \emph{might} and \emph{some}. The dependent measure in Noveck (2001) was the rate of ``logically correct responses'', i.e. responding ``yes'' to statements such as \emph{Some giraffes have long necks} or \emph{There might be a parrot [in the box]} when there had to be a parrot in the box. He found that children responded ``yes'' more frequently than adults, and concluded that children interpret scalar items \emph{some} and \emph{might} more logically (i.e. literally). Similarly in another landmark study, Papafragou and Musolino (2003) tested children and adults interpretation of the following set of scalar items: <\emph{two}, \emph{three}>, <\emph{some}, \emph{all}>, and <\emph{finish}, \emph{start}>. The dependent measure in this study was the proportion of ``No'' responses to a puppet's underinformative] statement. The study concluded that ``while adults overwhelmingly rejected infelicitous descriptions, children almost never did so''. Furthermore, the study compared  implicature rates across scales and concluded that ``children also differed from from adults in that their rejection rate on the numerical scale was reliably higher than on the two other scales.'' In their final experiment, Papafragou and Musolino (2003) modified their task to invite scalar inferences more easily. They reported that this manipulation resulted in a significantly higher rejection rates. Based on these results, they concluded that children's ability to compute implicatures is affected by the type of scalar item as well as children's awareness of the task's goals. Since these early pioneering studies, the rate of ``FALSE'' (or ``No'', ``Wrong'', ``Disagree'') responses on underinformative trials in truth-value judgment tasks has become a commonplace dependent measure (Doran et al (2012),  Geurts and Pouscoulous (2009), Potts, Lassiter, Levy, and Frank (2016), inter alia.)

Given the centrality of the theoretical notion of \enquote{implicature
rate} to much of experimental pragmatics, there is to date a surprising
lack of discussion of the basic assumption that it is adequately
captured by the proportion of ``FALSE' responses in truth value judgment
tasks (but see Benz and Gotzner (2014); Geurts and Pouscoulous (2009);
Degen and Goodman (2014); Katsos and Bishop (2011); Sikos et al (this issue)). Indeed, the scalar
implicature acquisition literature was shaken up when Katsos and Bishop
(2011) showed that simply by introducing an additional response option,
children started looking much more pragmatic than had been previously
observed in a binary judgment paradigm. Katsos and Bishop (2011) allowed
children to distribute a small, a big, or a huge strawberry to a puppet depending on
\enquote{how good the puppet said it}. The result was that children gave
on average smaller strawberries to the puppet when he produced
underinformative utterances compared to when he produced literally true
and pragmatically felicitous utterances, suggesting that children do, in
fact, display pragmatic ability even at ages when they had previously
appeared not to.

@@

We begin by reporting an experiment that addresses the following
question: do the number of response options provided in a truth value
judgment task and the way that responses are grouped into pragmatic
(\enquote{SI}) and literal (\enquote{no SI}) change inferences about
scalar implicature rates? Note that this way of asking the question
assumes two things: first, that whatever participants are doing in a
truth value judgment task, the behavioral measure can be interpreted as
providing a measure of interpretation; and second, that listeners either
do or do not compute an implicature on any given occasion. In the
General Discussion we will discuss both of these issues.
Following Degen and Goodman (2014), we will offer some remarks on why
truth value judgment tasks are better thought of as measuring
participants' estimates of speakers' \emph{production} probabilities.
This will suggest a completely different class of linking hypotheses.
We then discuss an alternative conception of scalar implicature as a
probabilistic phenomeonen, a view that has recently rose to prominence
in the subfield of probabilistic pragmatics (Franke \& Jäger, 2016;
Goodman \& Frank, 2016). This alternative conception of scalar
implicature, we argue, affords developing and testing quantitative
linking hypotheses in a rigorous and motivated way.

Consider a setup in which a listener is presented a card with a
depiction of either one or two animals (see Figure
\ref{fig:linkvisualization} for an example). As in a standard truth
value judgment task, the listener then observes an underinformative
utterance about this card (e.g., \enquote{There is a cat or a dog on the
card}) and is asked to provide a judgment on a scale with 2, 3, 4, or 5
response options, with endpoints \enquote{wrong} and
\enquote{right.}\footnote{An open question concerns the extent to which
  the labeling of points on the scale affects judgments (e.g.,
  \enquote{wrong}--\enquote{right} vs. \enquote{false}--\enquote{true}
  vs. \enquote{disagree}--\enquote{agree}). Studies vary in the labeling of scale points.} In the binary case, this reproduces the standard
truth value judgment task. Figure \ref{fig:linkvisualization}
exemplifies (some of) the researcher's options for grouping responses.
Under what we will call the \enquote{Strong link} assumption, only the
negative endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Under the \enquote{Weak link}
assumption, in contrast, any response that does not correspond to the
positive endpoint of the scale is interpreted as evidence for a scalar
implicature having been computed. Intermediate grouping schemes are also
possible, but these are the ones we will consider here. Note that for
the binary case, the Weak and Strong link return the same categorization
scheme, but for any number of response options greater than 2, the Weak
and Strong link can in principle lead to differences in inferences about
implicature rate.

@@

Our primary goal in this study was to test whether the estimated
implicature rate in the experimental task is affected by the linking
hypothesis and the number of response options available to participants.
To this end, we only analyzed the critical trials (exhaustive and
scalar). In particular, we classified each data point from critical
trials as constituting an implicature (1) or not (0) under the Strong
and Weak link. Figure \ref{fig:implicatureRatePlot} shows the resulting
implicature rates by condition and link. It is immediately apparent that there is variability in inferred implicature rate. In particular, the Weak link appears to result in greater estimates of
implicature rates in tasks with four or five response options, compared to the Strong link. For the binary and ternary task, the assumed link appears to play a much smaller role.

@@

The model provided evidence for the following effects: First, there was
a main effect of trial type such that scalar trials resulted in greater
implicature rates than exhaustive trials (Mean Estimate = 6.09, 95\%
Credible Interval={[}1, 12.29{]}). Second, there was an interaction
between link and number of response options such that the quaternary
task (Mean Estimate = 14.03, 95\% Credible Interval={[}7.24, 21.88{]})
and the quinary task (Mean Estimate = 17.28, 95\% Credible
Interval={[}10.64, 25.80{]}) resulted in greater
implicature rates with a weak link than with a strong link, but there was no evidence of a link-dependent difference in inferred implicature rate for the binary and ternary task. Finally, there was a three-way interaction between
link, trial type, and number of response options, driven by the binary/quinary contrast (Mean Estimate = -7.74,
95\% Credible Interval={[}-16.59, -0.16{]}). Simple effects analysis on only the binary and and quinary trials, separately for the exhaustive and scalar subset of the data, revealed that the three-way interaction is driven by a different effect of number of response options under the Weak vs Strong link for the two inference types. Specifically, on exhaustive trials, number of response options (2 vs.~5) only resulted in greater implicature rates under the Weak ($\beta$ = .2, $p <$ .0001), but not the Strong link ($\beta$ = -.8, $p <$ .82). In contrast, on scalar trials, number of response options (2 vs.~5) resulted in greater implicature rates under the Weak ($\beta$ = 3.6, $p <$ .005) link, and in lower implicature rates under the Strong link ($\beta$ = -4.0, $p <$ .0007).

In sum, both number of response options and link
affected the inferred implicature rate, as did the type of inference (exhaustive vs.~scalar).

@@

We take inspiration from the RSA approach and treat participants'
behavior in our experimental tasks as the result of a soft-optimal
pragmatic speaker in the RSA framework. That is, following Degen and
Goodman (2014), we proceed on the assumption that behavior on sentence
verification tasks such as truth value judgment tasks, is best modeled
as a function of an individual's mental representation of a cooperative
speaker (\(S_1\) in the language of RSA) rather than of a pragmatic
listener who interprets utterances (\(P_{L_1}\)).\footnote{Degen and Goodman (2014) argue that sentence verification is more plausibly construed as a production task rather than as an interpretation task because participants, unlike in natural language comprehension, are provided with the ground truth about the state of the world that a speaker is describing. Thus, participants are in essence being asked to assess the quality of a speaker's utterance. In contrast, Degen and Goodman argue, true interpretation tasks are characterized by the listener inferring what the state of the world is that the speaker is describing, for instance by selecting from one of multiple interpretation options.} In their paper, Degen
\& Goodman show that sentence verification tasks are relatively more
sensitive to contextual features like the
Question Under Discussion than are sentence interpretation tasks, and
that this follows if sentence interpretation tasks -- but not sentence
verification tasks -- require an additional layer of counterfactual
reasoning about the intentions of a cooperative speaker.

@@

The linking hypothesis under discussion assumes that speaker
probabilities of utterance given meaning are invariant across a) our
four different experimental conditions, b) across participants, and c)
within participants (that is, participants do not update
their \(S_1\) distribution in a local discourse context). We note that
the assumption (b) may conceivably be relaxed by allowing one or more of
the parameters in the model -- including the prior probability over
world states \(P_w\), the cost function on utterances \(C\), or the
rationality parameter \(\alpha\) -- to vary across participants. We also
note that assumption (c) in particular is in tension with a growing body
of empirical evidence that semantic and pragmatic interpretation is
modulated by rapid adaptation to the linguistic and social features of
one's interlocutors (Fine, Jaeger, Farmer, \& Qian, 2013; Kleinschmidt
\& Jaeger, 2015; Yildirim, Degen, Tanenhaus, \& Jaeger, 2016).

However, if we should like to keep the above simplifying assumptions in place, then
this linking hypothesis commits us to explaining variation in the data in terms
of the threshold parameters of our responder model \(R\). Consider first
the variation in response across different experimental conditions on a
given trial, e.g.~evaluation of a guess of \enquote{cat and dog} when
the card contains both a cat and a dog. The variation in the proportion
of responses of \enquote{right} on this trial between the binary,
ternary, quaternary, and quinary conditions indicates that the threshold
value for \enquote{right} responses must vary across conditions; that
is, we predict that the \(\theta\) of the binary condition will differ
from, e.g., the \(\theta_1\) of the ternary condition as well as the
\(\theta_1\) of the quaternary condition. We also observed variation in
response on this trial within a single condition (for example, a
sizeable minority of participants responded \enquote{wrong} to this
trial in the binary condition). Thus, this linking hypothesis is
committed to the notion that threshold values may vary across
participants, such that a speaker probability of utterance \(S_1\)(u
\textbar{} s) can fall below \(\theta\) for some subset of participants
while \(S_1\)(u \textbar{} s) itself remains constant across
participants.

Lastly, for two utterances of the same
conditional probability and in the same experimental condition, 
participants in our experiment sometimes provided a judgment of
\enquote{right} to one utterance but \enquote{wrong} to the other. That
is, there was within-subject variation in this
experiment. One way to represent such variation would be to posit that
the parameterization of threshold values proceeds stochastically and
that threshold values are recalibrated for every individual sentence
verification task. Rather than representing a threshold as a discrete
value N between 0 and 1, we can represent that threshold as a
distribution over possible threshold values -- with mass centered around
N. Whenever an individual encounters  a single trial of our truth value judgment task experiment, a
threshold value is sampled from this distribution. By allowing values of \(\theta\) to vary stochastically in this way, we can capture that \(S_1\)(u \textbar{} s) can fall both above and below \(\theta\) for a given
participant.

The model in its present form already captures an interesting asymmetry in inferred implicature rates between exhaustive and scalar trials of the experiment: note specifically (c.f. Figure 8) that inferred implicature rates are greater in the binary and ternary conditions for scalar trials over exhaustive trials. This is expected given the model's inferred speaker probabilities: the speaker probability of producing ``There is a cat on the card'' in the context of there being a cat and dog on the card (an exhaustive implicature-inducing trial) is greater than the speaker probability of producing ``There is a cat or a dog on the card'' in that same context (a scalar implicature-inducing trial). Assuming noisy  $\theta$ values centered around N, participants are expected to respond `Right' more frequently on exhaustive than on scalar trials, which is precisely what is observed. Recall that these probabilities were derived via the simplifying assumption of uniform cost on utterances; in fact, adding cost to relatively complex disjunctive sentences over simple declarative sentences only predicts a more pronounced asymmetry in the experimentally-observed direction.

@@

For the time being, however, we present the above analysis as a proof of
concept for the following idea: by relaxing the assumptions of the
traditional view of scalar implicature---namely, that scalar implicatures
either are or are not calculated, and that behavior on sentence
verification tasks directly reflects this binary interpretation
process---we can propose quantitative models of the variation in
behavior that is observed in experimental settings. We note that the linking
hypothesis proposed here is just one in the space of possible hypotheses. For
example, one might reject this threshold-based analysis in favor of one
whereby responses are the outcomes of sampling on the (pragmatic speaker
or pragmatic listener) probability distributions provided by an RSA
model. We leave this systematic, quantitative investigation to future work. For now we
emphasize that explicit computational modeling of behavioral responses is a tool that is  available to researchers in experimental
pragmatics. While using the RSA framework as the modeling tool requires revising traditional assumptions about the nature of scalar
implicature by relaxing the crisp notion of scalar
implicature as something that is or is not \enquote{calculated} in
interpretation, it provides new flexibility to explicitly discuss
 behavior in experimental settings. One need not adopt the RSA framework as the tool for hypothesizing and testing the link between theoretical constructs and behavior in pragmatic experiments. However, the empirical findings we have reported here---that the inferences researchers draw about ``implicature rate'' are volatile and depend on various features of the paradigm and the linking hypothesis employed--- strongly suggest that experimental pragmatics  as a field must engage more seriously with the foundational questions of what we are measuring in the experiments we run.

@@

\hypertarget{ref-Sikos2019}{}
Sikos, L., Kim, M., \& Grodner, D.J. (in press). Social context modulates tolerance for pragmatic violations in binary but not graded judgments. \emph{Frontiers in Psychology}.

@@

\hypertarget{ref-Yildirim2016}{}
Yildirim, I., Degen, J., Tanenhaus, M.K., \& Jaeger, T.F. (2016). Talker-specificity and adaptation in quantifier interpretation. \emph{Journal of Memory and Language}, \emph{87}, pp. 128-143.
